### 2013-01-24 ms.tex
commit rev 1

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex ms
% bibtex ms
% latex ms
% latex ms
% dvipdf ms

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
\usepackage{setspace}
\usepackage{hyperref} 
\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
%\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{The Case for Preprints in Biology}
}
% Insert Author names, affiliations and corresponding author email.
\\
Philippe Desjardins-Proulx$^{1, 2,\ast}$, 
Ethan P. White$^{3}$, 
Joel J. Adamson$^{4}$,
Karthik Ram$^{5}$,
Timoth\'ee Poisot$^{1,2,6}$,
Dominique Gravel$^{1,2}$
\\
\bf{1} Theoretical Ecosystem Ecology laboratory, Universit\'e du Qu\'ebec \`a Rimouski, Quebec, Canada.
\\
\bf{2} Quebec Center for Biodiversity Science, McGill University, Quebec, Canada.
\\
\bf{3} Departement of Biology, Utah State University, Utah, United-States of America.
\\
\bf{4} Ecology, Evolution and Organismic Biology, University of North Carolina at Chapel Hill, North Carolina, United-States of America.
\\
\bf{5} Environmental Science, Policy, and Management, University of California, Berkeley, California, United-States of America.
\\
\bf{6} International Network for Next-Generation Ecology.
\\
$\ast$ E-mail: philippe.d.proulx@gmail.com
\end{flushleft}

\section*{Introduction}

Public preprint servers allow authors to make manuscripts publicly available
before, or in parallel to, submitting them to journals for traditional peer-
review. The rationale for preprint servers is fundamentally simple: to make
the results of research available to the scientific community as
soon as possible, instead of waiting until the peer-review process is fully
completed.  Sharing manuscripts using preprint servers has numerous advantages
including: 1) rapid dissemination of work-in-progress to a wider audience; 2)
immediate visibility of the research output for early-career scientists; 3)
improved peer review by encouraging feedback from the entire research
community; and 4) a fair and straightforward way to establish precedence.

Open preprint servers offer a great opportunity for open science, especially if
the community embraces the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, can help make arXiv
attractive for scientists looking to promote their work \cite{lom12}. These
initiatives are important to fully exploit the potential of open preprint
servers. Posting preprints online increases the community of available informal
peer reviewers, and uses the internet for its original community-building
purposes.

Preprints began to gain popularity 20 years ago with the advent of arXiv, an
open preprint server widely used in physics and mathematics \cite{gin11}.
Preprints are also integral to the culture of other scientific fields.  Paul
Krugman noted that, in economics, the \emph{traditional model of submit, get
refereed, publish, and then people will read your work broke down a long time
ago. In fact, it had more or less fallen apart by the early 80s} \cite{kru12}.
In addition to a section on arXiv, economists have the RePEc (Research Papers in
Economics) initiative, which aims to create an archive of working papers,
manuscripts, and book chapters.

Despite the success of this approach in other fields, most manuscripts in
biology are not posted to preprint servers and are therefore not seen by more
than a handful of other scientists prior to publication. In this article, we: 1)
highlight the advantages of open preprint servers for both scientists and
publishers; 2) discuss the preprint policies of major publishers in biology; and
3) review popular options to submit preprints.

\section*{The case for public preprints}

The first and most often discussed advantage of open preprints is
speed (Figure~\ref{fig:map}). The time between submission and the official
publication of a manuscript can be measured in months, sometimes in years. For
all this time, the research is known only to a select few: colleagues, editors,
reviewers. Thus, the science cannot be used, discussed, or reviewed by the wider
scientific community. In a recent blog post, C. Titus Brown noted how posting a
paper on arXiv quickly led to a citation (arXiv papers can be cited) and his
research was used by another researcher \cite{bro12}. The current system of
hiding manuscripts before acceptance pose problems for both scientists and
publishers. Manuscripts that are unknown cannot be used and thus take more time
to be cited. It has been shown that high-energy physics, with its high arXiv
submission rate, had the highest immediacy among physics and mathematics
\cite{pra05}. Immediacy measures how quickly articles are cited.

Public preprints can be crucial to early-career scientists. The delay before
publication is seldom compatible with the pressure to show an impressive
publication record when applying for a scholarship or a position. Increasing the
perceived value of pre-prints as close, or equal, to journal articles will allow
young researchers to put their research outcome in the open, and build a
reputation for themselves through the diffusion of their work without fear that
this work will not be recognized by grant or job committees.

Posting manuscripts as preprints also has the potential to improve the 
quality of science by allowing prepublication feedback from a large pool of 
reviewers. In our experience, prepublication reviews by a small network of 
colleagues are common in the biological sciences and form an important part 
of the scientific process. These ``friendly'' reviews increases the chances 
of errors being caught prior to publication. Furthermore, the formal 
peer-review process as a whole is critically over-loaded. As the number of 
active scientists increases and the pressure to publish increases, it is 
becoming difficult for journals to find reviewers \cite{hoc09}.  At the same 
time, rejection rates are high in most journals \cite{aar08,roh09}, and when 
not invited to submit a revision, authors must start the process over again 
at another journal. As a result, initiatives to reduce time from submission 
to publication have emerged across the scientific community. Rohr et al.  
\cite{roh09} called for the recycling and reuse of peer-reviews: by 
attaching previous reviews and detailed replies to a new submission, both 
the editor and the referees can gauge the work done on the manuscript, and 
perhaps evaluate it with less prejudice. A widespread use of preprint 
servers can achieve the same goal of reducing the time spent in review. With 
a rich enough community of scientists depositing preprints, and commenting 
on them, the process of an open pre-review can become widespread and will 
overall increase the quality of first submissions \cite{hoc12}.

Finally, public preprint servers offer a fair way to establish intellectual
priority by making the work available as soon as it is complete. Some
manuscripts will spend much more time than others in the review process and/or in
production after acceptance. This means that publication and
acceptance dates do not accurately characterize who came up with an idea
first. For this reason, mathematicians and physicists have embraced arXiv in
part to establish priority in a fair way \cite{gin11,cal12}.

\section*{Preprints in biological sciences}

In contrast to other disciplines, the field of biology has effectively no
preprint culture, with the exception of small pockets of primarily highly
quantitative research (\emph{e.g.}, epidemiology, population genetics).  While
submitting to preprint servers has become more common in the past few years, the
number of biology papers submitted to preprint servers still represents only a
small fraction of the total research produced in biology (Fig.~\ref{arxiv}).

There are a number of reasons why biologists did not develop the reflex of
sharing preprints, many of which are based on common misconceptions. For
example, in contrast to other fields there is a perception in biology that
public preprints make it easier to steal ideas \cite{gin11}. In other fields
preprints serve the opposite role, they allow straightforward establishment of
precedence, letting research lay claim to an idea thus preventing it from being
``stolen'' \cite{gin11}

Another major concern is based on a certain interpretation of the Ingelfinger
rule: scientists should not publish the same manuscript twice \cite{alt96}. A
preprint is simply a document that allows ideas to spread and be discussed, it
is not yet formally validated by the peer-review system. This is why almost all
the major publishers in biology are preprint-friendly, including: Nature
Publishing Group, PLOS, BMC, PNAS, Elsevier, and Springer \ref{table:policies}.
This year, both the Ecological Society of America and the Genetics Society of
America changed their policies to allow public preprints. \emph{Nature} even
felt compelled to respond to the rumour that they refused manuscripts submitted
to arXiv by saying that ``\emph{Nature} never wishes to stand in the way of
communication between researchers. We seek rather to add value for authors and
the community at large in our peer review, selection and editing'' \cite{nat05}.
Still, a few journals adopt a ``by default'' hostile attitude towards preprints,
mostly due to the lack of clear policy of the publishers. As an example,
Wiley-Blackwell, which publishes some of the leading journals in biology, has no
official policy on the matter.

\section*{Current offerings}

We briefly discuss several options to submit preprints to open servers:
arXiv.org, figshare, and the upcoming PeerJ and F1000Research.

\subsection*{arXiv}

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in some branches of mathematics and physics.
arXiv provides a reliable citation system for all eprints. Physicist Paul
Ginsparg created arXiv in 1991 for theoretical high-energy physicists to
communicate preprints via email and ftp, and soon thereafter adopted the newly
created world-wide web \cite{jackson2002preprints}.  arXiv now receives over 7
000 submissions per month (\url{http://arxiv.org/show_monthly_submissions}) and
divides its submissions into subcategories of physics, mathematics, computer
science, quantitative biology, finance and statistics.  The quantitative biology
category includes subcategories for Populations and Evolution, Quantitative
Methods and other categories that may be of interest to biologists.

Unlike the other options presented here, arXiv has a system of moderators 
and endorsors. At least one author of a paper must be an endorser that has 
previously submitted a paper or has received permission to submit. 
Moderators have the power to change the classification of a paper. 
arXiv is now administered by the Cornell University Libraries, with funding 
coming from voluntary pledges by academic institutions along with matching 
funds from the Simons Foundation \cite{arxiv_future}. This approach to 
financing is one of the numerous measures that arXiv takes to ensure that 
the repository will remain permanently available and submissions will be 
readable.

\subsection*{figshare}

figshare (\href{http://figshare.com}{http://figshare.com}) is an open server
that allows scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output and, unlike arXiv, there is no endorser
system. All figshare content has a unique digital object identifier (DOI) like
any journal article, thus offering a permanent and stable link to the content.
A flexible tag system is used to classify each item. Comments can be made on all
content allowing for centralized discussion related to the material.

Unlike arXiv, figshare is not limited to quantitative sciences. arXiv.org has
sections on quantitative biology but might not be appropriate for
non-quantitative work. With its flexible approach to preprints, figshare offers
an alternative to arXiv for empirical biologists. Furthermore, by allowing all
types of content, figshare also provides an archive for early results
(\emph{e.g.}: figures, lab presentations).

\subsection*{PeerJ}

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new open access
publisher that combines both a preprint server and a peer reviewed journal.  It
is focused on the biological and medical sciences, which may help overcome the
perception that preprints do not have a place in biology. Like figshare, PeerJ
allows commenting on posted preprints, improving the potential for
pre-publication dialog. In addition, preprints can be made private if the
authors choose, and shared only with selected colleagues. While this reduces
many of the benefits of preprints described above, it may allow some researchers
who would not otherwise post preprints to begin exploring the possibility in a
manner appropriate to their current circumstances.

In contrast to other preprint servers, users cannot post unlimited public
preprints for free. One preprint per year can be posted for free, with a onetime
(\emph{i.e.} lifetime) fee of 99 dollars required to allow an author to post
unlimited public preprints. The preprint server is not tied to the journal, so
preprints can be posted regardless of where they will eventually be submitted
for publication.

\subsection*{F1000Research}\label{f1000r}

F1000Research is not a public preprint server like the previous three servers.
Whereas arXiv, figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually be
reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews that occur as part
of the submission process. Manuscripts are considered ``accepted'' and will only
be indexed after two positive referee responses. F1000Research works closely
with data providers to integrate raw data to the paper. For instance, upon
submitting a paper, authors are asked to upload their data, which are then
integrated in \emph{e.g.} figshare widgets, the DOI of which are given in the
paper when the data are first mentioned.  By integrating data to the paper,
F1000Research is working to make science more reproducible and open.

\subsection*{GitHub}

This manuscript was developed entirely as an open project on GitHub. GitHub is
one of several hosting services for collaborative development using the Git
version control system (VCS).  Git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. Git provides powerful features that allow numerous
contributers to work asynchronously on the same project, often in parallel
branches, all of which can be effortlessly merged and version controlled.  While
Git is created primarily for software development, where the use of version
control systems is standard \cite{aru12}, it is ideal for academic research
since it provides a way to collaborate on every step of the manuscript
development process. For example, during the development of this manuscript,
each author would clone the project (\emph{i.e.} make a personal copy), modify
it, and then merge their changes into a master branch. This takes the preprint
process to an new level, where the entire writing process is open from the
beginning.

\subsection*{Other options}

Scientific publishing is more diversified than ever and there are many
alternative options to submit articles before formal publications. 
Social networks such as ResearchGate can be used to submit preprints
\cite{lin12}. Also, if GitHub pushes openness further by opening the writing process,
open notebooks go even further by opening the entire scientific process
\cite{san11}.

\section*{Conclusion}

The ongoing discussions on the publication process, peer-reviewing and
alternative publication models are all symptoms of the current uneasiness
with the ever growing obsession with bibliographic metrics such as
the impact factor \cite{Fisher2012}. Researchers are pressured to orient
their publication strategy to maximize their number of publications and
total citations. A well-known consequence is to submit manuscripts first to the
most prestigious journals, and then resubmit to ``lower level'' journals as they
are rejected. The numerous negative impacts of such behavior have been discussed
in depth \cite{hoc09} and include a long delay between the time a manuscript is
finished to its publication. Research activities and the publication process
are drifting away from their fundamental object, namely the diffusion of novel
scientific discoveries.

Developing a preprint culture in biology will not solve all problems with the
current publication process. However, it might significantly reduce its negative
consequences. The role of peer-reviewing is to judge the scientific quality of a
study. It is the first barrier against the fraudulent and poor quality science
susceptible to impede scientific progress. In practice, the peer-review system
is not only used to evaluate scientific quality but also to judge pertinence. On
the other hand, preprints are not filtered, neither for their quality nor their
pertinence. Widespread adoption of preprint servers has the potential to shift
the diffusion strategy: journals would remain important to validate
publications, but the relevance of a study should
only be judged by many more readers than the typical two-four anonymous
reviewers. With a shift in the diffusion strategy, the role of traditional
journals and their editors would be to showcase scientific discoveries for
specialized readership.

Making publication easier can lead to the proliferation of studies of uneven
quality. A trade-off between the intensity of the peer-review filtering and the
benefits to science has been hypothesized \cite{Aarssen2012}.  With increasingly
stringent peer reviewing, the quality of published papers can improve at the
cost of censorship, an increased load on authors and reviewers, and greater
delays for publication.  Preprints are simply bypassing this model for what we
believe is the progress of science: they speed up the dissemination of
scientific discoveries and put on reader's shoulders the responsibility to judge
originality and pertinence.

\section*{Funding}

PDP is supported by an Alexander Graham Bell scholarship from the National
Sciences and Engineering Council of Canada. EPW is supported by a CAREER Award
from the National Science Foundation (DEB-0953694). JJA is supported by NSF
DEB-0614166 and NSF DEB-0919018.  TP is supported by a FQRNT-MELS post-doctoral
scholarship. KR is supported by NSF DEB-1021553. DG is funded by a Discovery
Grant from the National Sciences and Engineering Council of Canada and by the
Canada Research Chair program.

\section*{Acknowledgements}

We dedicate this article to Aaron H. Swartz (1986-2013).  We thank Carl
Boettiger, Mark Hahnel, and Hedvig Nenz\'en for helpful comments on an earlier
version of this manuscript.

\section*{References}
\bibliography{refs}

\newpage
\section*{Figure Legends}

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{map.eps}
\end{center}
\caption { {\bf It can take several months before a submitted paper is
officially published and citable.} Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are
given access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analyzed, and integrated
into current research. }
\label{fig:map}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{figure_2.png}
\end{center}
\caption { Submissions to the quantitative biology section lag behind
physics, mathematics, and computer science.
}
\label{fig:arxiv}
\end{figure}

\newpage
\section*{Tables}

\begin{table}[!ht]
    \caption{\bf{Policies for important publishers in biology.}}
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Genetics Society of America                 & Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Ecological Society of America       	& Accept \\
    Oxford Journals                             & Accept \\
    Science                             	& Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\
    British Ecological Society                  & No answer to our query \\
    \hline
    \end{tabular}
    \begin{flushleft} Some publishers tolerate preprints except for a few of
their medical journals, for example the \emph{ Journal of the National Cancer
Institute} from Oxford and \emph{The Lancet} from Elsevier.  \end{flushleft}
    \label{table:policies}
\end{table}

\end{document}